Shaping strategies and standardisation of faceted spheroids from Qianshangying (North China)

### 2025-06-14 analysis/paper/paper.qmd
updates

n_poly_multi <- group_counts$n[group_counts$Group == "Polyhedron + Multifacial"]
n_sub_spheroid <- group_counts$n[group_counts$Group == "Spheroid + Subspheroid"]
n_poly_multi

@@

p_value_pearson_sci <- pretty_print_sci(p_value)
```

@@

>>>>

n_poly_multi <- group_counts$n[group_counts$Group == "Polyhedron + Multifacial"]
n_sub_spheroid <- group_counts$n[group_counts$Group == "Spheroid + Subspheroid"]

@@

p_value_pearson_sci <- pretty_print_sci(p_value)

# 画图
ggplot(data_clean, aes(x = Sphericity, y = `SPHARM power`)) +
  # 散点（深红色）
  geom_point(size = 3, alpha = 0.9, color = "#B22222") +

  # 拟合线（深蓝灰）
  geom_smooth(method = "lm", se = TRUE, color = "#1C3041", fill = "#A3B1C6", linetype = "solid") +

   scale_x_continuous(breaks = seq(0.4, 1.0, by = 0.04)) +

  # 自定义配色主题
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_line(color = "#DDDDDD"),
    panel.grid.minor = element_blank(),
    axis.title = element_text(color = "#333333"),
    axis.text = element_text(color = "#333333")
  ) +

  # 标题与坐标轴标签
  labs(
    x = "Sphericity",
    y = "SPHARM Power"
  ) +
  annotate("text",
         x = Inf, y = Inf,
         hjust = 1.1, vjust = 1.2,
         label = paste0("r = ", round(correlation, 3),
                        "\n", "p = ", pretty_print_sci(p_value),
                        "\n", "n = ", nrow(data_clean)),
         size = 5, color = "#1C3041")
```

@@

```{r fig, echo=FALSE}

library(tidyverse)
library(ggrepel)

# read CSV 
umap_data <- read_csv(here("analysis/data/derived_data/umap_lmax20.csv"))

umap_data <- umap_data %>%
  mutate(category_lower = tolower(category)) %>%
  mutate(category_lower = fct_relevel(category_lower,
    "polyhedron", "multifacial", "idealmodel", "subspheroid", "spheroid"
  ))

# set colors
category_colors <- c(
  "idealmodel" = "black",
  "spheroid" = "red",
  "subspheroid" = "orange",
  "multifacial" = "lightgray",
  "polyhedron" = "gray40"
)

# point size
umap_data <- umap_data %>%
  mutate(point_size = if_else(category_lower == "idealmodel", 4.5, 3))

# label: idealmodel, spheroid, subspheroid
label_data <- umap_data %>%
  filter(category_lower %in% c("idealmodel", "spheroid", "subspheroid"))

# plot
ggplot(umap_data, aes(x = x, y = y)) +
  geom_point(aes(color = category_lower, size = point_size), alpha = 0.9, shape = 16) +
  geom_text_repel(
    data = label_data,
    aes(label = filename),
    size = 3,
    max.overlaps = 100,
    show.legend = FALSE
  ) +
  scale_color_manual(values = category_colors, name = "Category") +
  scale_size_identity() +
  theme_minimal(base_size = 14) +
  labs(
    x = "UMAP 1",
    y = "UMAP 2"
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )

```{r umap_cluster_highlight, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(here)

# 读取数据
umap_data <- read_csv(here("analysis/data/derived_data/umap_lmax20.csv"))

# KMeans 聚类（你可以根据 elbow method 先选 k）
set.seed(123)
kmeans_result <- kmeans(umap_data[, c("x", "y")], centers = 4)

# 将聚类结果加入数据框
umap_data <- umap_data %>%
  mutate(cluster = factor(kmeans_result$cluster))

# 可选：每个聚类颜色
cluster_colors <- RColorBrewer::brewer.pal(4, "Set2")

# 绘图：用 stat_ellipse 可视化每个簇的范围
ggplot(umap_data, aes(x = x, y = y, color = cluster)) +
  geom_point(alpha = 0.8, size = 3) +
  stat_ellipse(type = "norm", level = 0.95, linetype = "dashed") +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  theme_minimal(base_size = 14) +
  labs(
    title = "UMAP Projection with KMeans Clustering",
    x = "UMAP 1", y = "UMAP 2"
  )
```

```{r umap_cluster_highlight, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel)
library(here)

# Load data
umap_data <- read_csv(here("analysis/data/derived_data/umap_lmax20.csv"))

# KMeans clustering (k=4)
set.seed(123)
kmeans_result <- kmeans(umap_data[, c("x", "y")], centers = 4)

# Add cluster labels and idealmodel identifier
umap_data <- umap_data %>%
  mutate(
    cluster = factor(kmeans_result$cluster),
    is_idealmodel = if_else(category == "idealmodel", "yes", "no")
  )

# Color palette
cluster_colors <- RColorBrewer::brewer.pal(4, "Set2")

# Create the plot
ggplot(umap_data, aes(x = x, y = y)) +
  # Regular points (more transparent)
  geom_point(
    data = filter(umap_data, is_idealmodel == "no"),
    aes(color = cluster), 
    alpha = 0.8, 
    size = 2.5
  ) +
  # Highlight idealmodel points (darker, larger)
  geom_point(
    data = filter(umap_data, is_idealmodel == "yes"),
    aes(color = cluster),
    size = 3.5,
    show.legend = FALSE
  ) +
  # Add cluster ellipses
  stat_ellipse(
    aes(color = cluster),
    type = "norm", 
    level = 0.95, 
    linetype = "dashed",
    show.legend = FALSE
  ) +
  # Color and theme settings
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    x = "UMAP 1", 
    y = "UMAP 2",
    caption = "Dashed ellipses show 95% confidence intervals"
  ) +
  # Legend customization
  guides(
    color = guide_legend(override.aes = list(alpha = 1, size = 3))
  )
```
